---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
reticulate::py_require(
  "git+https://github.com/onnela-lab/forest@45fb41038bd46c25d9e6a4442aa74fa03b501317", 
  python_version = "3.11")
```

# walking

<!-- badges: start -->
[![R-CMD-check](https://github.com/jhuwit/walking/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/walking/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `walking` is to provide some algorithms to detect walking in tri-axial accelerometers.

## Installation

You can install the development version of walking from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("muschellij2/walking")
```

`walking` no longer tries to install `forest` for you. Install `forest` in
your own Python environment first, then load `walking` from R.

If you hit `clang++: error: unsupported option '-fopenmp'` or the LLVM
mismatch described in
[onnela-lab/forest#293](https://github.com/onnela-lab/forest/issues/293) and
[numba/llvmlite#1389](https://github.com/numba/llvmlite/issues/1389), use a
Python environment where `forest` and its dependencies already resolve cleanly,
then point `reticulate` at that environment with `use_condaenv()` or
`use_python()`.

### Parallel processing and `ssqueezepy`

`forest`'s `oak` workflow uses [`ssqueezepy`](https://github.com/OverLordGoldDragon/ssqueezepy#gpu--cpu-acceleration)
for wavelet processing. `ssqueezepy` can parallelize CPU work, which can make
larger analyses faster but may conflict with an R/Python environment or its
threading and OpenMP configuration. `find_walking()` disables this back-end
parallelization by default (`disable_parallelization = TRUE`) for a more
reliable R-session default. Set `disable_parallelization = FALSE` only when
you want to use the acceleration and your environment supports it.

If an error occurs while loading `forest`, or while running the wavelet step,
set `SSQ_PARALLEL` before the first `forest` import. This disables
`ssqueezepy` parallelization for the R session:

```r
Sys.setenv(SSQ_PARALLEL = 0)
walking::py_require_forest()
reticulate::import("forest")
```

See the [`ssqueezepy` GPU / CPU acceleration documentation](https://github.com/OverLordGoldDragon/ssqueezepy#gpu--cpu-acceleration)
for its parallelization settings.

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(walking)
csv_file = system.file("test_data_bout.csv", package = "walking")
x = readr::read_csv(csv_file)
colnames(x)[colnames(x) == "UTC time"] = "time"
res = find_walking(data = x)
print(res)
```

## Potential Conflicts

### Running `forest` and [`stepcount`](https://github.com/jhuwit/stepcount). 
The two Python modules (`forest` and `stepcount`) can have different conflicts of python dependencies and python versions.  You can do specific call-outs with `callr` to try to resolve this:

## Running Walking with Different Python Environments

If you want to use both walking estimation from `forest` and `stepcount`, you can run them in separate R processes using `callr` to avoid Python package conflicts. 
Here's how you can do it:

```{r, cache = TRUE}
library(callr)
library(stepcount)
csv_file = system.file("test_data_bout.csv", package = "walking")
data = readr::read_csv(csv_file)
colnames(data)[colnames(data) == "UTC time"] = "time"

stepcount_callr = function(data,
                           ...) {
  
  reticulate::py_require("stepcount==3.11.0", python_version = "3.10")
  sc <- reticulate::import("stepcount")
  stepcount::stepcount_check()
  
  res = stepcount::stepcount(data, ...)
  return(res)
}

# 2. Run the isolated background R process
result <- callr::r(
  func = stepcount_callr,
  show = TRUE,
  args = list(data = data) # Safely injects data into the process
)
head(result)

forest_callr = function(data,
                        ...) {
  reticulate::py_require(
    "git+https://github.com/onnela-lab/forest@45fb41038bd46c25d9e6a4442aa74fa03b501317", 
    python_version = "3.11")
  fr = reticulate::import("forest")
  oak = fr$oak$base
  oak
  res = walking::estimate_steps_forest(data, ...)
  return(res)
}

# 2. Run the isolated background R process
fresult <- callr::r(
  func = forest_callr,
  show = TRUE,
  args = list(data = data) # Safely injects data into the process
)
head(fresult)
```
